---
title: "prepareDensityandCause"
output: html_document
---
   Written by: Jessica Yu, 2022.                                                                         
               jessica_yu@alumni.ubc.ca                                                                               
                                                                                             
   Data from: Statistics Canada, Population Data BC, BC Ministry of Health, and The Canadian Urban Environmental Health Research Consortium    
   Paper under review from this work: 
   Yu J, Gustafson P, Tran M, Brauer M. Assessing trade-offs and optimal ranges of densities for life expectancy and 12 causes of mortality in Metro Vancouver, British Columbia, Canada, 1990-2016. International Journal of Environmental Research and Public Health. 2022.                                          
                                                                                                         
   **Outline**                                                                                               
             (A) Load libraries, import data, and define attributes                  
             (B) Load all density files and merge with cause files
             (C) OPTION 1: Density by log scale
             (D) OPTION 2: Density by absolute scale
             (E) OPTION 3: Density by absolute scale for all years
             (F) OPTION 4: Density by SES group
             (G) OPTION 5: Cubic spline stratification by SES

#Load libraries, import data, and define attributes    
            
```{r}
rm(list=ls())
library(dplyr)
library(ggplot2)
library(car)
dir<-"R:/working/modeling/temp_dir/causes/"
dir2<-"R:/working/preparing data/"
dir3<-"R:/working/modeling/cov_files/"
dir4<-"R:/working/results/density/causes/"
dir5<-"R:/working/results/ses/"
this_cause<-"diab_ckd"
file<-".Rdata"
file2<-".pdf"
CTs<-368
load(paste0(dir3, "covar_", CTs,"CTs", file))
load("R:/working/modeling/cov_files/education_368CTs.RData")
education<-education %>% rename(area=CT)

```

## Load all density files and merge with cause files - change to cause of choice

```{r}

year<-c(1990:2016)


for (i in year) {
  load(paste0(dir,"sims/",this_cause, "/mx_est_", i,file)) ##CHANGE TO CAUSE OF INTEREST
  est <- est %>% filter(age==99) %>% select(year,area, mx_mean) 
  assign(paste0("birth",i), est)
  rm(est)
}

birthAll<-lapply(ls(pattern="birth"), get)
library(plyr)
birthAll<-ldply(birthAll, data.frame)
detach("package:plyr", unload=TRUE)


load("R:/working/preparing data/Density/popdensity_final_368.RDATA")
load(paste0(dir2, "CT", CTs, file))

densityAll<-merge(density,CT, by="area")

birthAll<-left_join(birthAll, densityAll, by=c("area","year"))

```

## OPTION 1: Density by log scale

```{r}
##ALL DATA
dir2<-"R:/working/preparing data/"
file<-".Rdata"
this_cause<-"intent"

load(paste0(dir2, "density/birthAll_",this_cause, file))

pdf(paste0(dir4,this_cause,"/Density_",this_cause,"_all",file2), width=12, height=12)

densityPlot<- 
  ggplot(birthAll3, aes(x=log10(population_density), y=mx_mean, label=msdgroup, color=msdgroup)) +
  geom_smooth(method=lm, formula=y ~ splines::bs(x,3),  color="red", fill="#69b3a2", se=TRUE) 
densityPlot + labs(x="Population Density (Log Transformed)", y=paste0(this_cause, " Mortality (per 100,000)"), colour="SES", title=paste0("Population Density and ",this_cause," Mortality"),  cex.main=2, cex.sub=1.5, cex.sub=1.25)

dev.off()

##HIGH SES DATA

birthAll4<-birthAll3 %>% filter(msdgroup==5)

pdf(paste0(dir4,this_cause,"/Density_",this_cause,"_highSES",file2), width=12, height=12)

densityPlot<- 
  ggplot(birthAll4, aes(x=log10(population_density), y=mx_mean, label=msdgroup, color=msdgroup)) +
  geom_point() +
  geom_smooth(method=lm, formula=y ~ splines::bs(x,3),  color="red", fill="#69b3a2", se=TRUE) 
densityPlot + labs(x="Population Density (Log Transformed)", y=paste0(this_cause, " Mortality (per 100,000)"), colour="SES", title=paste0("Population Density and ",this_cause," Mortality"), cex.main=2, cex.sub=1.5, cex.sub=1.25)

dev.off()

birthAll4<-birthAll3 %>% filter(msdgroup==1)

pdf(paste0(dir4,this_cause,"/Density_",this_cause,"_lowSES",file2), width=12, height=12)

densityPlot<- 
  ggplot(birthAll4, aes(x=log10(population_density), y=mx_mean, label=msdgroup, color=msdgroup)) +
  geom_point() +
  geom_smooth(method=lm, formula=y ~ splines::bs(x,3),  color="red", fill="#69b3a2", se=TRUE) 
densityPlot + labs(x="Population Density (Log Transformed)", y=paste0(this_cause, " Mortality (per 100,000)"), colour="SES", title=paste0("Population Density and ",this_cause," Mortality"), cex.main=2, cex.sub=1.5, cex.sub=1.25)

dev.off()

```


## OPTION 2: Density by absolute scale


```{r}
dir4<-"R:/working/results/density/causes/"
dir2<-"R:/working/preparing data/"
file<-".Rdata"
this_cause<-"intent"

load(paste0(dir2, "density/birthAll_",this_cause, file))


pdf(paste0(dir4,this_cause,"/Density_",this_cause,"_all_absolute",file2), width=12, height=12)

densityPlot<- 
  ggplot(birthAll3, aes(x=population_density, y=mx_mean, label=msdgroup, color=msdgroup)) +
  geom_point() +
  geom_smooth(method=lm, formula=y ~ splines::bs(x,3),  color="red", fill="#69b3a2", se=TRUE) 
densityPlot + labs(x="Population Density", y=paste0(this_cause, " Mortality (per 100,000)"), colour="SES", title=paste0("Population Density and ",this_cause," Mortality"),  cex.main=2, cex.sub=1.5, cex.sub=1.25) 

dev.off()

##HIGH SES DATA

birthAll4<-birthAll3 %>% filter(msdgroup==5)

pdf(paste0(dir4,this_cause,"/Density_",this_cause,"_highSES_absolute",file2), width=12, height=12)

densityPlot<- 
  ggplot(birthAll4, aes(x=population_density, y=mx_mean, label=msdgroup, color=msdgroup)) +
  geom_point() +
  geom_smooth(method=lm, formula=y ~ splines::bs(x,3),  color="red", fill="#69b3a2", se=TRUE) 
densityPlot + labs(x="Population Density", y=paste0(this_cause, " Mortality (per 100,000)"), colour="SES", title=paste0("Population Density and ",this_cause," Mortality"), cex.main=2, cex.sub=1.5, cex.sub=1.25)

dev.off()

birthAll4<-birthAll3 %>% filter(msdgroup==1)

pdf(paste0(dir4,this_cause,"/Density_",this_cause,"_lowSES_absolute",file2), width=12, height=12)

densityPlot<- 
  ggplot(birthAll4, aes(x=population_density, y=mx_mean, label=msdgroup, color=msdgroup)) +
  geom_point() +
  geom_smooth(method=lm, formula=y ~ splines::bs(x,3),  color="red", fill="#69b3a2", se=TRUE) 
densityPlot + labs(x="Population Density", y=paste0(this_cause, " Mortality (per 100,000)"), colour="SES", title=paste0("Population Density and ",this_cause," Mortality"), cex.main=2, cex.sub=1.5, cex.sub=1.25)

dev.off()

```


## OPTION 3: Density by absolute scale for all years


```{r}
##ALL DATA
this_cause<-"hiv_std"

load(paste0(dir2, "density/birthAll_",this_cause, file))

pdf(paste0(dir4,this_cause,"/Density_",this_cause,"_all_absolute",file2), width=12, height=12)

densityPlot<- 
  ggplot(birthAll3, aes(x=population_density, y=mx_mean, label=msdgroup, color=msdgroup)) +
  geom_smooth(method=lm, formula=y ~ splines::bs(x,3),  color="red", fill="#69b3a2", se=TRUE) 
densityPlot + labs(x="Population Density", y=paste0(this_cause, " Mortality (per 100,000)"), colour="SES", title=paste0("Population Density and ",this_cause," Mortality"), cex.main=2, cex.sub=1.5, cex.sub=1.25)

dev.off()

```


## OPTION 4: Density by SES group

```{r}
load(paste0(dir2, "density/birthAll_",this_cause, file))
cause<-"Maternal and neonatal diseases"
year<-2016

birthAll3<-birthAll3 %>% filter(year==year)

#SES only
png(paste0(dir5,this_cause,"_ses_",year,".png"))

ggplot(birthAll3, aes(factor(msdgroup), mx_mean)) + 
  xlab("SES")+
  ylab("Mortality Rate per 100,000") +
  ggtitle(paste0(cause))+
  geom_bar(stat="identity", position = "dodge", fill="#f68060") + 
  scale_fill_brewer(palette = "Oranges") 
dev.off()

png(paste0(dir5,this_cause,"_ses_density_",year,".png"))
#SES and density
ggplot(birthAll3, aes(factor(msdgroup), mx_mean, fill = densitygroup)) + 
  xlab("SES")+
  ylab("Mortality Rate per 100,000") +
  labs(fill="Density") +
  ggtitle(paste0(cause))+
  geom_bar(stat="identity", position = "dodge") + 
  scale_fill_brewer(palette = "Oranges") 
dev.off()

```

## OPTION 5: Cubic spline stratification by SES

```{r}

load("R:/working/preparing data/density/birthAll_final_368.RDATA")
birthAll2<-birthAll %>% merge(CTCSD, by="area") %>% select(area,year,mx_mean,ex_mean,population_density,CSDNAME)
names(covar)[2]<-"area"



birthAll2<-birthAll2 %>% left_join(covar, by=c("year","area")) 

outliers<-Boxplot(ex_mean ~ cut(population_density, 10), data=birthAll2, main="Outliers of Population Density and LE for All Years", cex.axis=0.5, id.method="expdiff_percentage")

birthAll3<-birthAll2 %>%
  filter(!row_number() %in% outliers)



birthAll2<-birthAll2 %>% group_by(year) %>% mutate(msdgroup=ntile(msd, 5), matgroup=ntile(MatScore, 5), socgroup=ntile(SocScore, 5), edugroup=ntile(education, 5))

matscore<-ggplot(birthAll3, aes(x=population_density, y=ex_mean, color=as.factor(matgroup))) +
  geom_smooth(data=birthAll3, aes(x=population_density, y=ex_mean, color=as.factor(matgroup))) +
  labs(x="Population Density", y="Life Expectancy", color="Material Deprivation Score", title="Population Density and Life Expectancy", cex.main=2, cex.sub=1.5, cex.sub=1.25) +
  theme_bw()

  
socscore<-ggplot(birthAll3, aes(x=population_density, y=ex_mean, color=as.factor(socgroup))) +
  geom_smooth(data=birthAll3, aes(x=population_density, y=ex_mean, color=as.factor(socgroup)))  +
  labs(x="Population Density", y="Life Expectancy", color="Social Deprivation Score", title="Population Density and Life Expectancy", cex.main=2, cex.sub=1.5, cex.sub=1.25) +
  theme_bw()  
  
msdscore<-ggplot(birthAll3, aes(x=population_density, y=ex_mean, color=as.factor(msdgroup))) +
  geom_smooth(data=birthAll3, aes(x=population_density, y=ex_mean, color=as.factor(msdgroup)))  +
  labs(x="Population Density", y="Life Expectancy", color="Material and Social Deprivation Score", title="Population Density and Life Expectancy", cex.main=2, cex.sub=1.5, cex.sub=1.25) +
  theme_bw()



```

