---
title: "prepareDerivative"
output: html_document
---

   Written by: Jessica Yu, 2022.                                                                         
               jessica_yu@alumni.ubc.ca                                                                               
                                                                                             
   Data from: Statistics Canada, Population Data BC, BC Ministry of Health, and The Canadian Urban Environmental Health Research Consortium    
   Paper under review from this work: 
   Yu J, Gustafson P, Tran M, Brauer M. Assessing trade-offs and optimal ranges of densities for life expectancy and 12 causes of mortality in Metro Vancouver, British Columbia, Canada, 1990-2016. International Journal of Environmental Research and Public Health. 2022.                                          
                                                                                                         
   **Outline**                                                                                               
             (A) Load libraries, import data, and define attributes                                                  
             (B) Apply cubic splines for density and cause-specific mortality rates and find derivative
             (C) Find inflection and plot derivative function
             (D) Plot new cause-specific mortality rate figures with shaded area for optimal range
             (E) Apply cubic splines for density and life expectancy and find derivative
             (F) Plot new life expectancy figures with shaded area for optimal range
             

#Load libraries, import data, and define attributes 

```{r}
library(ggplot2)
library(car)
library(dplyr)
this_cause<-"neo"
dir2<-"R:/working/preparing data/"
dir4<-"R:/working/results/density/causes/"
file<-".Rdata"
load(paste0(dir2, "density/birthAll_",this_cause, file))
```


## Apply cubic splines for density and cause-specific mortality rates and find derivative

```{r}

birthAll3 <-birthAll3 %>% filter(year==2016)

model= lm(birthAll3$mx_mean ~ splines::bs(birthAll3$population_density,3))
X<-data.frame(birthAll3$population_density)
Y<-predict(model, newdata=X)
X<-X[,1]
plot(X, Y, type="p", main="Original fit")

dY<-diff(Y)/diff(X)
dX<-rowMeans(embed(X, 2))
plot(dX, dY, type="p", main="Derivative")

original<-data.frame(X,Y)
fitted<-data.frame(dX,as.data.frame(dY))


```

## Find inflection and plot derivative function

```{r}
#Inflection min if U-shaped derivative function or max if inverted U shape
inflection<-fitted[which.max(fitted$dY),] #min if u-shaped
lowest<-original$X[which.min(original$Y)]

pdf(paste0("R:/working/results/density/causes/", this_cause, "/", this_cause, "_derivative.pdf"), width=12, height=12)

ggplot(fitted, aes(x=dX, y=dY)) +
  ggtitle(paste("Derivative of", this_cause, "function")) +
  geom_point(y=dY, colour="red") + 
  theme(text=element_text(size=30), legend.title=element_text(size=40)) +
  stat_smooth() +
  geom_vline(xintercept=inflection[1,1], colour="grey", linetype="dotdash", size=1.5)  

dev.off()

inflection<-inflection[1,1]

shade<-data.frame(xmin=lowest, xmax=inflection, ymin=-Inf, ymax=Inf)
```


## Plot new cause-specific mortality rate figures with shaded area for optimal range

```{r}
pdf(paste0(dir4,this_cause,"/Density_",this_cause,"_all_absolute_range_2016.pdf"), width=12, height=12)

densityPlot<- 
  ggplot(birthAll3, aes(x=population_density, y=mx_mean, label=msdgroup, color=msdgroup)) +
  #ylim(0.000255,0.00028) +
  geom_point() +
  theme(text=element_text(size=30), legend.title=element_text(size=30), legend.text = element_text(size=30)) +
  #facet_grid(CSDNAME ~ year) +
  geom_smooth(method=lm, formula=y ~ splines::bs(x,3),  color="red", fill="#69b3a2", se=TRUE) 
densityPlot + labs(x="Population Density", y=paste0("Mortality Rate (per 100,000)"), colour="SES", title=paste0(this_cause),  cex.main=2, cex.sub=1.5, cex.sub=1.25) +
  geom_vline(xintercept=inflection, colour="grey", linetype="dotdash", size=1.5) +
  geom_vline(xintercept=lowest, colour="grey", linetype="dotdash", size=1.5) +
  geom_rect(data=shade, aes(xmin=xmin, xmax=xmax, ymin=ymin, ymax=ymax), fill="yellow", alpha=0.1, inherit.aes=FALSE)

dev.off()
```


## Apply cubic splines for density and life expectancy and find derivative

```{r}

load("R:/working/preparing data/density/birthAll_final_368_withcovar.RDATA")

outliers<-Boxplot(ex_mean ~ cut(population_density, 10), data=birthAll2, main="Outliers of Population Density and LE for All Years", cex.axis=0.5, id.method="expdiff_percentage")

birthAll2<-birthAll2 %>%
  filter(!row_number() %in% outliers) %>%
  filter(year==2016)

model= lm(birthAll2$ex_mean ~ splines::bs(birthAll2$population_density,3))

X<-data.frame(birthAll2$population_density)
Y<-predict(model, newdata=X)
X<-X[,1]
plot(X, Y, type="p", main="Original fit")

dY<-diff(Y)/diff(X)
dX<-rowMeans(embed(X, 2))
plot(dX, dY, type="p", main="Derivative")

original<-data.frame(X,Y)
fitted<-data.frame(dX,as.data.frame(dY))

inflection<-fitted[which.max(fitted$dY),]
lowest<-original$X[which.max(original$Y)]

pdf(paste0("R:/working/results/density/LE/LE_derivative_alllSES_2016.pdf"), width=12, height=12)

ggplot(fitted, aes(x=dX, y=dY)) +
  ggtitle(paste("Derivative of LE function")) +
  geom_point(y=dY, colour="red") +
  theme(text=element_text(size=30), legend.title=element_text(size=40)) + 
  stat_smooth() +
  geom_vline(xintercept=inflection[1,1], colour="grey")  

dev.off()

inflection<-inflection[1,1]
shade<-data.frame(xmin=lowest, xmax=inflection, ymin=-Inf, ymax=Inf)

```

##  Plot new life expectancy figures with shaded area for optimal range

```{r}
pdf(paste0("R:/working/results/density/LE/DensityandLe_allSES_range_2016.pdf"), width=12, height=12)

densityPlot<- 
  ggplot(birthAll2, aes(x=population_density, y=ex_mean, label=msdgroup, color=msdgroup)) +
  geom_point() +
  theme(text=element_text(size=40)) +
  geom_smooth(method=lm, formula=y ~ splines::bs(x,3),  color="red", fill="#69b3a2", se=TRUE) 
densityPlot + labs(x="Population Density", y="Life Expectancy", colour="SES", title="Population Density and Life Expectancy", cex.main=2, cex.sub=1.5, cex.sub=1.25)+
  geom_vline(xintercept=inflection, colour="grey", linetype="dotdash", size=1.5) +
  geom_vline(xintercept=lowest, colour="grey", linetype="dotdash", size=1.5) +
  geom_rect(data=shade, aes(xmin=xmin, xmax=xmax, ymin=ymin, ymax=ymax), fill="yellow", alpha=0.1, inherit.aes=FALSE)

dev.off()
```


