►---
title: "prepareDensity"
output: html_document
---

   Written by: Jessica Yu, 2022.                                                                         
               jessica_yu@alumni.ubc.ca                                                                               
                                                                                             
   Data from: Statistics Canada, Population Data BC, BC Ministry of Health, and The Canadian Urban Environmental Health Research Consortium    
   Paper under review from this work: 
   Yu J, Gustafson P, Tran M, Brauer M. Assessing trade-offs and optimal ranges of densities for life expectancy and 12 causes of mortality in Metro Vancouver, British Columbia, Canada, 1990-2016. International Journal of Environmental Research and Public Health. 2022.                                          
                                                                                                         
   **Outline**                                                                                               
             (A) Load libraries, import data, and define attributes                                                  
             (B) Load all LE estimates
             (C) Load all density files
             (D) OPTION 1: Cross sectional plot of all years
             (E) OPTION 2: Analyse with absolute scale
             (F) OPTION 3: Analyse by SES groups 
             (G) OPTION 4: Analyse by SES groups - 1991-2016
             (H) OPTION 5: Analyse by SES groups, all years, and municipality
             
## Load libraries, import data, and define attributes 

```{r}
rm(list=ls())
library(dplyr)
library(ggplot2)
library(car)

dir<-"R:/working/modeling/temp_dir/"
dir2<-"R:/working/preparing data/"
dir3<-"R:/working/modeling/cov_files/"
dir4<-"R:/working/results/density/LE/"

CTs<-368
file<-".rdata"

load(paste0(dir2, "CT", CTs, file))
CTCSD<-read.csv(file="R:/working/CTCSD.csv")
CTCSD<-left_join(CT,CTCSD, by="CT")
load(paste0(dir3, "covar_", CTs,"CTs", file))
load(paste0(dir3, "covar_with_edu", CTs,"CTs", file))


```

## Load all LE estimates

```{r}

dir<-"R:/working/modeling/temp_dir/"
year<-1990:2016

for (i in year) {
  load(paste0(dir, "lt_est_", i, ".rdata"))
  assign(paste0("birth",i), est)
  rm(est)
}

birthAll<-lapply(ls(pattern = "birth"), get)
birthAll<-lapply(birthAll, function(x) subset(x, sex==3 & age==0))
birthAll<-lapply(birthAll, function(x) select(x, "area", "year","sex","mx_mean","ex_mean"))

birthAll2<-do.call(rbind.data.frame, birthAll)

  
```


## Load all density files

```{r}
load("R:/working/preparing data/Density/popdensity_final_368.RDATA")

densityAll<-density

birthAll3<-merge(birthAll2,densityAll, by=c("area","year"))

birthAll<-birthAll3
save(birthAll, file="R:/working/preparing data/density/birthAll_final_368.RDATA")
```


## OPTION 1: Cross sectional plot of all years - facet for years and CSD, and identify SES groups

```{r}
load("R:/working/preparing data/density/birthAll_final_368.RDATA")
birthAll2<-birthAll %>% merge(CTCSD, by="area") %>% select(area,year,mx_mean,ex_mean,population_density,CSDNAME)
names(covar)[2]<-"area"

birthAll2<-birthAll2 %>% left_join(covar, by=c("year","area")) 

birthAll2<-birthAll2 %>% group_by(year) %>% mutate(MatScore=1-MatScore, SocScore=1-SocScore, msd=1-msd) %>% mutate(msdgroup=ntile(msd, 5), matgroup=ntile(MatScore, 5), socgroup=ntile(SocScore, 5), edugroup=ntile(education, 5))

save(birthAll2, file="R:/working/preparing data/density/birthAll_final_368_withcovar.RDATA")


outliers<-Boxplot(ex_mean ~ cut(population_density, 10), data=birthAll2, main="Outliers of Population Density and LE for All Years", cex.axis=0.5, id.method="expdiff_percentage")

birthAll3<-birthAll2 %>%
  filter(!row_number() %in% outliers)


pdf(paste0(dir4, "DensityandLe_all_absolute.pdf"), width=12, height=12)

densityPlot<- 
  ggplot(birthAll3, aes(x=population_density, y=ex_mean, label=msdgroup, color=msdgroup)) +
  geom_point() +
  geom_smooth(method=lm, formula=y ~ splines::bs(x,3),  color="red", fill="#69b3a2", se=TRUE) 
densityPlot + labs(x="Population Density (Log Transformed)", y="Life Expectancy", colour="SES", title="Population Density and Life Expectancy", subtitle="Using Census and Log Transformed Mortality data from 1990-2016 and outliers removed", cex.main=2, cex.sub=1.5, cex.sub=1.25)

dev.off()

birthAll3<-filter(birthAll3, year %in% c(1991,2001,2016))
birthAll4<-filter(birthAll3, CSDNAME %in% c("Vancouver","Burnaby","Surrey","Richmond","North Vancouver"))
birthAll4.low<-filter(birthAll4, msdgroup==1)
birthAll4.high<-filter(birthAll4, msdgroup==5)


birthAll5<-filter(birthAll3, CSDNAME %in% c("Coquitlam","New Westminster","Langley","Maple Ridge","West Vancouver"))

#Look at Vancouver, Richmond, North Vancouver, Burnaby, and Surrey

pdf(paste0(dir4, "DensityandLe_group1", ".pdf"), width=12, height=12)

densityPlot<- 
  ggplot(birthAll4, aes(x=log10(population_density), y=ex_mean)) +
  geom_point() +
  facet_grid(CSDNAME ~ year) +
  geom_smooth(method=lm, formula=y ~ splines::bs(x,3),  color="red", fill="#69b3a2", se=TRUE) 
densityPlot + labs(x="Population Density (Log Transformed)", y="Life Expectancy", colour="Material and Social Deprivation Index", title="Population Density and Life Expectancy", subtitle="Using Census and Mortality data from 1991-2016", cex.main=2, cex.sub=1.5, cex.sub=1.25)

dev.off()


pdf(paste0(dir4, "DensityandLe_group2", ".pdf"), width=12, height=12)

densityPlot<- 
  ggplot(birthAll5, aes(x=log10(population_density), y=ex_mean, label=msdgroup, color=msdgroup)) +
  geom_point() +
  facet_grid(CSDNAME ~ year) +
  geom_smooth(method=lm, formula=y ~ splines::bs(x,3),  color="red", fill="#69b3a2", se=TRUE) 
densityPlot + labs(x="Population Density (Log Transformed)", y="Life Expectancy", colour="Material and Social Deprivation Index", title="Population Density and Life Expectancy", subtitle="Using Census and Mortality data from 1991-2016", cex.main=2, cex.sub=1.5, cex.sub=1.25)

dev.off()


```

## OPTION 2: Analyse with absolute scale

```{r}
load("R:/working/preparing data/density/birthAll_final_368.RDATA")
birthAll2<-birthAll %>% merge(CTCSD, by="area") %>% select(area,year,mx_mean,ex_mean,population_density,CSDNAME)
names(covar)[2]<-"area"

birthAll2<-birthAll2 %>% left_join(covar, by=c("year","area")) 

birthAll2<-birthAll2 %>% group_by(year) %>% mutate(MatScore=1-MatScore, SocScore=1-SocScore, msd=1-msd) %>% mutate(msdgroup=ntile(msd, 5), matgroup=ntile(MatScore, 5), socgroup=ntile(SocScore, 5), edugroup=ntile(education, 5))

save(birthAll2, file="R:/working/preparing data/density/birthAll_final_368_withcovar.RDATA")

outliers<-Boxplot(ex_mean ~ cut(population_density, 10), data=birthAll2, main="Outliers of Population Density and LE for All Years", cex.axis=0.5, id.method="expdiff_percentage")

birthAll3<-birthAll2 %>%
  filter(!row_number() %in% outliers)


pdf(paste0(dir4, "DensityandLe_all_absolute.pdf"), width=12, height=12)

densityPlot<- 
  ggplot(birthAll3, aes(x=population_density, y=ex_mean, label=matgroup, color=matgroup)) +
  geom_point() +
  geom_smooth(method=lm, formula=y ~ splines::bs(x,3),  color="red", fill="#69b3a2", se=TRUE) 
densityPlot + labs(x="Population Density", y="Life Expectancy", colour="SES", title="Population Density and Life Expectancy", cex.main=2, cex.sub=1.5, cex.sub=1.25)

dev.off()

birthAll4.low<-filter(birthAll3, msdgroup==1)
birthAll4.high<-filter(birthAll3, msdgroup==5)

pdf(paste0(dir4, "DensityandLe_low_absolute", ".pdf"), width=12, height=12)

densityPlot<- 
  ggplot(birthAll4.low, aes(x=population_density, y=ex_mean)) +
  geom_point() +
  geom_smooth(method=lm, formula=y ~ splines::bs(x,3),  color="red", fill="#69b3a2", se=TRUE) 
densityPlot + labs(x="Population Density", y="Life Expectancy", colour="SES", title="Population Density and Life Expectancy - Low SES", cex.main=2, cex.sub=1.5, cex.sub=1.25)

dev.off()

#High income

pdf(paste0(dir4, "DensityandLe_high_absolute", ".pdf"), width=12, height=12)

densityPlot<- 
  ggplot(birthAll4.high, aes(x=population_density, y=ex_mean)) +
  geom_point() +
  geom_smooth(method=lm, formula=y ~ splines::bs(x,3),  color="red", fill="#69b3a2", se=TRUE) 
densityPlot + labs(x="Population Density", y="Life Expectancy", colour="SES", title="Population Density and Life Expectancy - High SES", cex.main=2, cex.sub=1.5, cex.sub=1.25)

dev.off()
```


## OPTION 3: Analyse by SES groups - 2016 only

```{r}
load("R:/working/preparing data/density/birthAll_final_368.RDATA")
birthAll2<-merge(birthAll, CTCSD, by="area")
names(covar)[2]<-"area"
birthAll2<-left_join(birthAll2,covar, by=c("year","area"))

birthAll2$msdgroup<-.bincode(birthAll2$msd,
                                breaks = quantile(birthAll2$msd, seq(0,1, by=0.2)),
                                include.lowest = TRUE)


outliers<-Boxplot(ex_mean ~ cut(population_density, 10), data=birthAll2, main="Outliers of Population Density and LE for All Years", cex.axis=0.5, id.method="expdiff_percentage")

birthAll3<-birthAll2 %>%
  filter(!row_number() %in% outliers)

birthAll3<-filter(birthAll3, year %in% c(2016))
birthAll4.low<-filter(birthAll3, msdgroup==1)
birthAll4.high<-filter(birthAll3, msdgroup==5)
#Low income

pdf(paste0(dir4, "DensityandLe_group1_low", ".pdf"), width=12, height=12)

densityPlot<- 
  ggplot(birthAll4.low, aes(x=log10(population_density), y=ex_mean)) +
  geom_point() +
  geom_smooth(method=lm, formula=y ~ splines::bs(x,3),  color="red", fill="#69b3a2", se=TRUE) 
densityPlot + labs(x="Population Density (Log Transformed)", y="Life Expectancy", colour="Material and Social Deprivation Index", title="Population Density and Life Expectancy - Low Material and Social Deprivation Index Scores", subtitle="Using Census and Mortality data from 2016", cex.main=2, cex.sub=1.5, cex.sub=1.25)

dev.off()

#High income

pdf(paste0(dir4, "DensityandLe_group1_high", ".pdf"), width=12, height=12)

densityPlot<- 
  ggplot(birthAll4.high, aes(x=log10(population_density), y=ex_mean)) +
  geom_point() +
  geom_smooth(method=lm, formula=y ~ splines::bs(x,3),  color="red", fill="#69b3a2", se=TRUE) 
densityPlot + labs(x="Population Density (Log Transformed)", y="Life Expectancy", colour="Material and Social Deprivation Index", title="Population Density and Life Expectancy - High Material and Social Deprivation Index Scores", subtitle="Using Census and Mortality data from 2016", cex.main=2, cex.sub=1.5, cex.sub=1.25)

dev.off()

```


## OPTION 4: Analyse by SES groups - 1991-2016

```{r}
load("R:/working/preparing data/density/birthAll_final_368.RDATA")
birthAll2<-merge(birthAll, CTCSD, by="area")
names(covar)[2]<-"area"
birthAll2<-left_join(birthAll2,covar, by=c("year","area"))

birthAll2$msdgroup<-.bincode(birthAll2$msd,
                                breaks = quantile(birthAll2$msd, seq(0,1, by=0.2)),
                                include.lowest = TRUE)


outliers<-Boxplot(ex_mean ~ cut(population_density, 10), data=birthAll2, main="Outliers of Population Density and LE for All Years", cex.axis=0.5, id.method="expdiff_percentage")

birthAll3<-birthAll2 %>%
  filter(!row_number() %in% outliers)

birthAll4.low<-filter(birthAll3, msdgroup==1)
birthAll4.high<-filter(birthAll3, msdgroup==5)
#Low income

pdf(paste0(dir4, "DensityandLe_group1_low_allyears", ".pdf"), width=12, height=12)

densityPlot<- 
  ggplot(birthAll4.low, aes(x=log10(population_density), y=ex_mean)) +
  geom_point() +
  geom_smooth(method=lm, formula=y ~ splines::bs(x,3),  color="red", fill="#69b3a2", se=TRUE) 
densityPlot + labs(x="Population Density (Log Transformed)", y="Life Expectancy", colour="Material and Social Deprivation Index", title="Population Density and Life Expectancy - Low Material and Social Deprivation Index Scores", subtitle="Using Census and Mortality data from 1991-2016", cex.main=2, cex.sub=1.5, cex.sub=1.25)

dev.off()

#High income

pdf(paste0(dir4, "DensityandLe_group1_high_allyears", ".pdf"), width=12, height=12)

densityPlot<- 
  ggplot(birthAll4.high, aes(x=log10(population_density), y=ex_mean)) +
  geom_point() +
  geom_smooth(method=lm, formula=y ~ splines::bs(x,3),  color="red", fill="#69b3a2", se=TRUE) 
densityPlot + labs(x="Population Density (Log Transformed)", y="Life Expectancy", colour="Material and Social Deprivation Index", title="Population Density and Life Expectancy - High Material and Social Deprivation Index Scores", subtitle="Using Census and Mortality data from 1991-2016", cex.main=2, cex.sub=1.5, cex.sub=1.25)

dev.off()

```


## OPTION 5: Analyse by SES groups, all years, and municipality

```{r}
load("R:/working/preparing data/density/birthAll_final_368.RDATA")
birthAll2<-merge(birthAll, CTCSD, by="area")
names(covar)[2]<-"area"
birthAll2<-left_join(birthAll2,covar, by=c("year","area"))

birthAll2$msdgroup<-.bincode(birthAll2$msd,
                                breaks = quantile(birthAll2$msd, seq(0,1, by=0.2)),
                                include.lowest = TRUE)


outliers<-Boxplot(ex_mean ~ cut(population_density, 10), data=birthAll2, main="Outliers of Population Density and LE for All Years", cex.axis=0.5, id.method="expdiff_percentage")

birthAll3<-birthAll2 %>%
  filter(!row_number() %in% outliers)

birthAll4<-filter(birthAll3, CSDNAME %in% c("Langley"))


birthAll4.low<-filter(birthAll4, msdgroup==1)
birthAll4.high<-filter(birthAll4, msdgroup==5)
#Low income

pdf(paste0(dir4, "DensityandLe_Langley_low_allyears", ".pdf"), width=12, height=12)

densityPlot<- 
  ggplot(birthAll4.low, aes(x=log10(population_density), y=ex_mean)) +
  geom_point() +
  geom_smooth(method=lm, formula=y ~ splines::bs(x,3),  color="red", fill="#69b3a2", se=TRUE) 
densityPlot + labs(x="Population Density (Log Transformed)", y="Life Expectancy", colour="Material and Social Deprivation Index", title="Population Density and Life Expectancy - Low Material and Social Deprivation Index Scores", subtitle="Using Census and Mortality data from 1991-2016", cex.main=2, cex.sub=1.5, cex.sub=1.25)

dev.off()

#High income

pdf(paste0(dir4, "DensityandLe_Langley_high_allyears", ".pdf"), width=12, height=12)

densityPlot<- 
  ggplot(birthAll4.high, aes(x=log10(population_density), y=ex_mean)) +
  geom_point() +
  geom_smooth(method=lm, formula=y ~ splines::bs(x,3),  color="red", fill="#69b3a2", se=TRUE) 
densityPlot + labs(x="Population Density (Log Transformed)", y="Life Expectancy", colour="Material and Social Deprivation Index", title="Population Density and Life Expectancy - High Material and Social Deprivation Index Scores", subtitle="Using Census and Mortality data from 1991-2016", cex.main=2, cex.sub=1.5, cex.sub=1.25)

dev.off()

```
