---
title: "prepareDID"
output: html_document
---

   Written by: Jessica Yu, 2022.                                                                         
               jessica_yu@alumni.ubc.ca                                                                               
                                                                                             
   Data from: Statistics Canada, Population Data BC, BC Ministry of Health, and The Canadian Urban Environmental Health Research Consortium    
   Paper under review from this work: 
   Yu J, Gustafson P, Tran M, Brauer M. Assessing trade-offs and optimal ranges of densities for life expectancy and 12 causes of mortality in Metro Vancouver, British Columbia, Canada, 1990-2016. International Journal of Environmental Research and Public Health. 2022.                                          
                                                                                                         
   **Outline**                                                                                               
             (A) Load libraries, import data, and define attributes                                                  
             (B) Merge data and assess the change in density from 1990 to 2016
             (C) Visualise the density change  
             (D) Plot density plot of density change to use as threshold decision
             (E) Plot parallel line assumption - absolute thresholds
             (F) Merge density change and prepare DID dataset
             (G) Apply DID model by percent change
             (H) Create shapefile for treated and control identification
             (I) Apply DID model on LE for High SES group
             (J) Apply DID model on LE for Low SES Group
             (K) Apply DID model on cause-specific mortalities - percent change
             (L) Apply DID model on cause-specific mortalities - low SES group
             (M) Apply DID model on cause-specific mortalities - high SES group
             (N) Descriptive data for paper
             (O) Derive land area

## Load libraries, import data, and define attributes  

```{r}
rm(list=ls())
library(dplyr)
library(ggplot2)
library(rgdal)
load("R:/working/preparing data/density/birthAll_final_368.RDATA") ##LIFE EXPECTANCY WITH DENSITY FOR ALL CENSUS YEARS
load("R:/working/preparing data/CT368.rdata")
CTs<-1:368
```

## Merge data and assess the change in density from 1990 to 2016

```{r}
#initialise a dataframe
densityChange<-data.frame("area" = 1:368, "change" = 0)
birthAll$area<-birthAll$area+1

#loop to look at differences in density from 1991 to 2016

for (i in CTs) {
  densityChange[i,2]<-birthAll$population_density[which((birthAll$year==2016) & (birthAll$area==i))]-birthAll$population_density[which((birthAll$year==1990) & (birthAll$area==i))]
}


#check the densities in 1991 and 2016 for manual inspection

densityChangeCheck<-data.frame("area" = 1:368,"density1991"=0,"density2016"=0, "change" = 0)

for (i in CTs) {
  densityChangeCheck[i,2]<-birthAll$population_density[which((birthAll$year==1990) & (birthAll$area==i))]
  densityChangeCheck[i,3]<-birthAll$population_density[which((birthAll$year==2016) & (birthAll$area==i))]
  densityChangeCheck[i,4]<-birthAll$population_density[which((birthAll$year==2016) & (birthAll$area==i))]-birthAll$population_density[which((birthAll$year==1990) & (birthAll$area==i))]
}

birthAll$area<-birthAll$area-1
densityChangeCheck$area<-densityChangeCheck$area-1

densityChangeCheck<-merge(densityChangeCheck,CT, by="area")
densityChangeCheck<-densityChangeCheck %>%
  mutate(percent_change=change/density1991)

save(densityChangeCheck, file="R:/working/preparing data/density/densityChangeCheck.RDATA")
write.csv(densityChangeCheck, file="R:/working/preparing data/density/densityChangeCheck.csv", row.names = F)

```

## Visualise the density change  

```{r}
load(file="R:/working/modeling/shape_files/shape-mapping368CTs.RData")
shape$CTNAME<-as.numeric(as.character(shape$CTNAME))


range<-round(quantile(densityChange$change, c(.05,.1,.25,.5,.75,.9,.95), na.rm=TRUE))
at=c(range)

labels=c(paste0(""),range)


#Prep LE
densityChanges<-densityChange
names(densityChanges)[1]<-"CTNAME"

densityChangeCheck <- densityChangeCheck %>% rename(CTNAME=area)

#merge shapefiles with current Exp and Mortality
shapeChange<-merge(x=shape, y=densityChangeCheck, by="CTNAME")

writeOGR(shapeChange, dsn = "DensityChange", layer="densityChange", driver="ESRI Shapefile")

pdf("R:/working/preparing data/Density/Figures/densityChangesMap.pdf")

names(shapeChange@data)[2]<-"change"
shapeChange$change<-as.numeric(shapeChange$change)
shapeChange@data$change<-cut(shapeChange@data$change, breaks=c(min(densityChange$change),at, max(densityChange$change)))
spplot(shapeChange, "change",  sp.layout=list("sp.polygons",shapeChange),par.settings = list(axis.line=list(col="transparent")), main="Density Changes from 1991 to 2016",colorkey=list(labels=list(at=seq(0.5,length(at)+1), labels=labels)))

dev.off()

```


## Plot density plot of density change to use as threshold decision

```{r}
library(tidyverse)

quantile(densityChangeCheck$change, .25) #71.3

`pdf("R:/working/preparing data/Density/Figures/histogramDensity.pdf") 

p1<- densityChange %>%
  ggplot(aes(x=change)) +
  geom_histogram(binwidth=368, fill="#69b3a2", color="#e9ecef", alpha=0.9) +
  ggtitle("Histogram of Changes in Population Density (1991-2016)") 
  theme(
    plot.title = element_text(size=15))
  
p1
  
p2<- densityChange %>%
  filter ( change<6000) %>%
  ggplot(aes(x=change)) +
  geom_histogram(binwidth=368, fill="#69b3a2", color="#e9ecef", alpha=0.9) +
  ggtitle("Histogram of Changes in Population Density (1991-2016)") 
  theme(
    plot.title = element_text(size=15))
  
p2

dev.off()  
`
```


## Plot parallel line assumption - absolute thresholds


```{r}

load("R:/working/preparing data/density/birthAll_final_368_withcovar.RDATA")
load("R:/working/preparing data/density/densityChangeCheck.RDATA")

birthAll3<-merge(birthAll2,densityChangeCheck,by="area")


birthall<-birthAll3
#birthall<-subset(birthall, birthall$change>0)

x.95th<-quantile(birthall$change, c(.95))
x.90th<-quantile(birthall$change, c(.90))
x.75th<-quantile(birthall$change, c(.75))
x.50th<-quantile(birthall$change, c(.5))


#birthall<-subset(birthall, change<0)#80th percentile and those less than 0
birthall<-subset(birthall, change>=quantile(birthall$change, c(.90)) | change<=quantile(birthall$change, c(.10)))#90th and 10th percentile


#Find density change data 
birthall = birthall %>%
  mutate(treated = change>=x.90th)


#Plot year and life expectancy of two density change categories to ensure parallel assumption is met
pdf("R:/working/preparing data/Density/Figures/parallel_line_assumption_90v10alldata.pdf") 

ggplot(birthall, aes(year, ex_mean, color = treated)) +
  stat_summary(geom = 'line') +
  geom_vline(xintercept = 1996) +
  ggtitle("Parallel Line Assumption for Densification of P90 vs P10 - All CTs")+
  xlab("Year") +
  ylab("Life Expectancy") +
  labs(color="Densification") +
  scale_color_manual(labels= c("P10", "P90"), values=c("blue","red"))+
  theme_minimal()

dev.off()

```

#Merge change and prepare DID dataset

```{r}

load("R:/working/preparing data/density/birthAll_final_368_withcovar.RDATA")
load("R:/working/preparing data/density/densityChangeCheck.RDATA")

birthall<-birthAll

birthall<-merge(birthall,densityChange,by="area")
birthall<-subset(birthall, birthall$change>0)

birthall<-subset(birthall, change>=quantile(birthall$change, c(.90)) | change<=quantile(birthall$change, c(.10)))

x.95th<-quantile(birthall$change, .95)
x.90th<-quantile(birthall$change, .90)
x.80th<-quantile(birthall$change, .80)
x.75th<-quantile(birthall$change, .75)
x.50th<-quantile(birthall$change, .50)




birthall$time = ifelse(birthall$year>1996,1,0)
birthall$treated = ifelse(birthall$change>=x.90th,1,0)
birthall$did = birthall$time * birthall$treated

didreg = lm(mx_mean ~ treated + time + did, dat = birthall)

sink("R:/working/preparing data/Density/Figures/didsummary_90v10.txt") 

print(summary(didreg)) 

sink()

```


## Apply DID model by percent change

```{r}
load("R:/working/preparing data/density/birthAll_final_368_withcovar.RDATA")
#Or for summary stats use this so we keep mat and soc deprivation scores in tact
#load("R:/working/preparing data/density/birthAll_final_368_withcovarmsd.RDATA")

load("R:/working/preparing data/density/densityChangeCheck.RDATA")

#for LE
birthAll3<-merge(birthAll2,densityChangeCheck,by="area")


birthall<-birthAll3


x.90th<-quantile(birthall$percent_change, c(.90))
x.80th<-quantile(birthall$percent_change, c(.80))
x.75th<-quantile(birthall$percent_change, c(.75))
x.50th<-quantile(birthall$percent_change, c(.50))

#birthall<-subset(birthall, change<0)#80th percentile and those less than 0
birthall<-subset(birthall, percent_change>=quantile(birthall$percent_change, c(.90)) | percent_change<=quantile(birthall$percent_change, c(.10)))#95th and 5th percentile


#Find density change data 
birthall = birthall %>%
  mutate(treated = percent_change>=x.90th)

#Plot year and life expectancy of two density change categories to ensure parallel assumption is met
pdf("R:/working/preparing data/Density/Figures/parallel_line_assumption_90v10alldata.pdf") 

ggplot(birthall, aes(year, ex_mean, color = treated)) +
  stat_summary(geom = 'line') +
  geom_vline(xintercept = 1996) +
  ggtitle("Parallel Line Assumption for Densification of P90 vs P10 - All CTs")+
  xlab("Year") +
  ylab("Life Expectancy") +
  labs(color="Densification") +
  scale_color_manual(labels= c("P10", "P90"), values=c("blue","red"))+
  theme_minimal()

dev.off()

birthall$time = ifelse(birthall$year>1996,1,0)
birthall$treated = ifelse(birthall$percent_change>=x.90th,1,0)
birthall$did = birthall$time * birthall$treated

didreg = lm(mx_mean ~ treated + time + did, dat = birthall)
CIs<-confint(didreg, "did", level=0.95)
CIs_LE<-confint(didreg, "did", level=0.95)


sink("R:/working/preparing data/Density/Figures/didsummary_alldata_90v10.txt") 

print(summary(didreg)) 
print(CIs)

sink()

birthall<-birthAll3

#birthall<-subset(birthall, change<0)#80th percentile and those less than 0
birthall<-subset(birthall, change>=quantile(birthall$percent_change, c(.75)) | change<=quantile(birthall$percent_change, c(.25)))#95th and 5th percentile

#Find density change data 
birthall = birthall %>%
  mutate(treated = percent_change>=x.75th)

#Plot year and life expectancy of two density change categories to ensure parallel assumption is met
pdf("R:/working/preparing data/Density/Figures/parallel_line_assumption_75v25alldata.pdf") 

ggplot(birthall, aes(year, ex_mean, color = treated)) +
  stat_summary(geom = 'line') +
  geom_vline(xintercept = 1996) +
  ggtitle("Parallel Line Assumption for Densification of P75 vs P25 - All CTs")+
  xlab("Year") +
  ylab("Life Expectancy") +
  labs(color="Densification") +
  scale_color_manual(labels= c("P10", "P90"), values=c("blue","red"))+
  theme_minimal()

dev.off()

birthall$time = ifelse(birthall$year>1996,1,0)
birthall$treated = ifelse(birthall$percent_change>=x.75th,1,0)
birthall$did = birthall$time * birthall$treated

didreg = lm(ex_mean ~ treated + time + did, dat = birthall)

sink("R:/working/preparing data/Density/Figures/didsummary_alldata_75v25.txt") 

print(summary(didreg)) 

sink()


birthall<-birthAll3

#birthall<-subset(birthall, change<0)#80th percentile and those less than 0
birthall<-subset(birthall, change>=quantile(birthall$percent_change, c(.50)) | change<quantile(birthall$percent_change, c(.50)))#95th and 5th percentile

#Find density change data 
birthall = birthall %>%
  mutate(treated = percent_change>=x.50th)

#Plot year and life expectancy of two density change categories to ensure parallel assumption is met
pdf("R:/working/preparing data/Density/Figures/parallel_line_assumption_50v50alldata.pdf") 

ggplot(birthall, aes(year, ex_mean, color = treated)) +
  stat_summary(geom = 'line') +
  geom_vline(xintercept = 1996) +
  ggtitle("Parallel Line Assumption for Densification of P50 vs P50 - All CTs")+
  xlab("Year") +
  ylab("Life Expectancy") +
  labs(color="Densification") +
  scale_color_manual(labels= c("P10", "P90"), values=c("blue","red"))+
  theme_minimal()

dev.off()

birthall$time = ifelse(birthall$year>1996,1,0)
birthall$treated = ifelse(birthall$percent_change>=x.50th,1,0)
birthall$did = birthall$time * birthall$treated

didreg = lm(ex_mean ~ treated + time + did, dat = birthall)

sink("R:/working/preparing data/Density/Figures/didsummary_alldata_50v50.txt") 

print(summary(didreg)) 

sink()

```

## Create shapefile for treated and control identification

```{r}
load(file="R:/working/modeling/shape_files/shape-mapping368CTs.RData")
shape$CTNAME<-as.numeric(as.character(shape$CTNAME))

treated <- birthall %>%
  select(area, treated) %>%
  group_by(area) %>%
  summarise(treated=median(treated)) %>%
  rename(CTNAME=area) %>%
  mutate(treated=treated+1)

shapeTreated<-merge(x=shape, y=treated, by="CTNAME")

writeOGR(shapeTreated, dsn = "DensityChange_lowSES", layer="densityTreated_lowSES", driver="ESRI Shapefile", overwrite_layer = T)

```


## Apply DID model on LE for High SES group

```{r}
load("R:/working/preparing data/density/birthAll_final_368_withcovar.RDATA")
#Or for summary stats use this so we keep mat and soc deprivation scores in tact
#load("R:/working/preparing data/density/birthAll_final_368_withcovarmsd.RDATA")

load("R:/working/preparing data/density/densityChangeCheck.RDATA")
birthAll2<-merge(birthAll2,densityChangeCheck,by="area")

birthall<-birthAll2 %>% filter(msdgroup==1)

#Only use below when calculating descript because we need to subset year
#birthall<-birthAll2 %>% filter(msdgroup==1 & year<1995)


#birthall<-birthAll2 %>% filter(incomegroup==5)


x.95th<-quantile(birthall$percent_change, c(.95))
x.90th<-quantile(birthall$percent_change, c(.90))
x.80th<-quantile(birthall$percent_change, c(.80))
x.75th<-quantile(birthall$percent_change, c(.75))
x.50th<-quantile(birthall$percent_change, c(.5))

birthall<-subset(birthall, percent_change>=quantile(birthall$percent_change, c(.90)) | percent_change<=quantile(birthall$percent_change, c(.10)))

birthall = birthall %>%
  mutate(treated = percent_change>=x.90th)

allCTsHigh<-data.frame(unique(birthall$area))


pdf("R:/working/preparing data/Density/Figures/parallel_line_assumption_90v10_highSESdata.pdf") 

ggplot(birthall, aes(year, ex_mean, color = treated)) +
  stat_summary(geom = 'line') +
  geom_vline(xintercept = 1996) +
  ggtitle("Parallel Line Assumption Check for Upper 90th vs Bottom 10th percentile (1990-2016) - Low SES Data")+
  theme_minimal()

dev.off()

#DID

birthall$time = ifelse(birthall$year>1996,1,0)
birthall$treated = ifelse(birthall$percent_change>=x.90th,1,0)
birthall$did = birthall$time * birthall$treated

didreg = lm(ex_mean ~ treated + time + did, dat = birthall)
CIs_LE_high<-confint(didreg, "did", level=0.95)

sink("R:/working/preparing data/Density/Figures/didsummary_highSES_90v10.txt") 

print(summary(didreg)) 

sink()


birthall<-birthAll2 %>% filter(msdgroup==1)

#birthall<-subset(birthall, change<0)#80th percentile and those less than 0
birthall<-subset(birthall, change>=quantile(birthall$percent_change, c(.75)) | change<=quantile(birthall$percent_change, c(.25)))#95th and 5th percentile


#Find density change data 
birthall = birthall %>%
  mutate(treated = percent_change>=x.75th)

#Plot year and life expectancy of two density change categories to ensure parallel assumption is met
pdf("R:/working/preparing data/Density/Figures/parallel_line_assumption_75v25_highSES.pdf") 

ggplot(birthall, aes(year, ex_mean, color = treated)) +
  stat_summary(geom = 'line') +
  geom_vline(xintercept = 1996) +
  ggtitle("Parallel Line Assumption for Densification of P75 vs P25 - High SES CTs")+
  xlab("Year") +
  ylab("Life Expectancy") +
  labs(color="Densification") +
  scale_color_manual(labels= c("P10", "P90"), values=c("blue","red"))+
  theme_minimal()

dev.off()

birthall$time = ifelse(birthall$year>1996,1,0)
birthall$treated = ifelse(birthall$percent_change>=x.75th,1,0)
birthall$did = birthall$time * birthall$treated

didreg = lm(ex_mean ~ treated + time + did, dat = birthall)

sink("R:/working/preparing data/Density/Figures/didsummary_75v25_highSES.txt") 

print(summary(didreg)) 

sink()

```

## Apply DID model on LE for Low SES Group

```{r}
load("R:/working/preparing data/density/birthAll_final_368_withcovar.RDATA")
#Or for summary stats use this so we keep mat and soc deprivation scores in tact
#load("R:/working/preparing data/density/birthAll_final_368_withcovarmsd.RDATA")

load("R:/working/preparing data/density/densityChangeCheck.RDATA")
birthAll2<-merge(birthAll2,densityChangeCheck,by="area")

#birthall<-birthAll2 %>% filter(msdgroup==5)

birthall<-birthAll2 %>% filter(msdgroup==5)
#birthall<-birthAll2 %>% filter(incomegroup==5)


x.95th<-quantile(birthall$percent_change, c(.95))
x.90th<-quantile(birthall$percent_change, c(.90))
x.80th<-quantile(birthall$percent_change, c(.80))
x.75th<-quantile(birthall$percent_change, c(.75))
x.50th<-quantile(birthall$percent_change, c(.5))

birthall<-subset(birthall, percent_change>=quantile(birthall$percent_change, c(.90)) | percent_change<=quantile(birthall$percent_change, c(.10)))#95th and 5th percentile

birthall = birthall %>%
  mutate(treated = percent_change>=x.90th)

allCTsLow<-data.frame(unique(birthall$area))


pdf("R:/working/preparing data/Density/Figures/parallel_line_assumption_90v10_lowSESdata.pdf") 

ggplot(birthall, aes(year, ex_mean, color = treated)) +
  stat_summary(geom = 'line') +
  geom_vline(xintercept = 1996) +
  ggtitle("Parallel Line Assumption Check for Upper 75th vs Bottom 25th percentile (1990-2016) - High SES Data")+
  theme_minimal()

dev.off()

#DID

birthall$time = ifelse(birthall$year>1996,1,0)
birthall$treated = ifelse(birthall$percent_change>=x.90th,1,0)
birthall$did = birthall$time * birthall$treated

didreg = lm(ex_mean ~ treated + time + did, dat = birthall)
CIs_LE_low<-confint(didreg, "did", level=0.95)

sink("R:/working/preparing data/Density/Figures/didsummary_lowSES_90v10.txt") 

print(summary(didreg)) 

sink()

birthall<-birthAll3

#birthall<-subset(birthall, change<0)#80th percentile and those less than 0
birthall<-subset(birthall, change>=quantile(birthall$percent_change, c(.75)) | change<=quantile(birthall$percent_change, c(.25)))#95th and 5th percentile


#Find density change data 
birthall = birthall %>%
  mutate(treated = percent_change>=x.75th)

#Plot year and life expectancy of two density change categories to ensure parallel assumption is met
pdf("R:/working/preparing data/Density/Figures/parallel_line_assumption_75v25_lowSES.pdf") 

ggplot(birthall, aes(year, ex_mean, color = treated)) +
  stat_summary(geom = 'line') +
  geom_vline(xintercept = 1996) +
  ggtitle("Parallel Line Assumption for Densification of P75 vs P25 - Low SES CTs")+
  xlab("Year") +
  ylab("Life Expectancy") +
  labs(color="Densification") +
  scale_color_manual(labels= c("P10", "P90"), values=c("blue","red"))+
  theme_minimal()

dev.off()

birthall$time = ifelse(birthall$year>1996,1,0)
birthall$treated = ifelse(birthall$percent_change>=x.75th,1,0)
birthall$did = birthall$time * birthall$treated

didreg = lm(ex_mean ~ treated + time + did, dat = birthall)

sink("R:/working/preparing data/Density/Figures/didsummary_75v25_lowSES.txt") 

print(summary(didreg)) 

sink()

```


## Apply DID model on cause-specific mortalities - percent change


```{r}
this_cause<-"cvd"
dir2<-"R:/working/preparing data/"
file<-".Rdata"
load(paste0(dir2, "density/birthAll_",this_cause, file))

load("R:/working/preparing data/density/densityChangeCheck.RDATA")


birthAll3<-merge(birthAll3,densityChangeCheck,by="area")

birthall<-birthAll3

x.90th<-quantile(birthall$percent_change, c(.90))
x.75th<-quantile(birthall$percent_change, c(.75))


birthall<-subset(birthall, percent_change>=quantile(birthall$percent_change, c(.90)) | percent_change<=quantile(birthall$percent_change, c(.1)))

#Find density change data 
birthall = birthall %>%
  mutate(treated = percent_change>=x.90th)

#Plot year and life expectancy of two density change categories to ensure parallel assumption is met
pdf(paste0("R:/working/preparing data/Density/Figures/parallel_line_assumption_90v10_", this_cause, ".pdf")) 

ggplot(birthall, aes(year, mx_mean, color = treated)) +
  stat_summary(geom = 'line') +
  geom_vline(xintercept = 1996) +
  ggtitle("Parallel Line Assumption for Densification of P90 vs P10 - All CTs")+
  xlab("Year") +
  ylab("Life Expectancy") +
  labs(color="Densification") +
  scale_color_manual(labels= c("P10", "P90"), values=c("blue","red"))+
  theme_minimal()

dev.off()


#DID

birthall$time = ifelse(birthall$year>1996,1,0)
birthall$treated = ifelse(birthall$percent_change>=x.90th,1,0)
birthall$did = birthall$time * birthall$treated

didreg = lm(mx_mean ~ treated + time + did, dat = birthall)
CIs<-confint(didreg, "did", level=0.95)

assign(paste0("CIs_",this_cause), CIs)

sink(paste0("R:/working/preparing data/Density/Figures/didsummary_90v10_", this_cause, ".txt")) 

print(summary(didreg)) 

sink()


#75 vs 25
birthall<-birthAll3

birthall<-subset(birthall, percent_change>=quantile(birthall$percent_change, c(.75)) | percent_change<=quantile(birthall$percent_change, c(.25)))

#Find density change data 
birthall = birthall %>%
  mutate(treated = percent_change>=x.75th)

#Plot year and life expectancy of two density change categories to ensure parallel assumption is met
pdf(paste0("R:/working/preparing data/Density/Figures/parallel_line_assumption_75v25_", this_cause, ".pdf")) 

ggplot(birthall, aes(year, mx_mean, color = treated)) +
  stat_summary(geom = 'line') +
  geom_vline(xintercept = 1996) +
  ggtitle("Parallel Line Assumption for Densification of P75 vs P25 - All CTs")+
  xlab("Year") +
  ylab("Life Expectancy") +
  labs(color="Densification") +
  scale_color_manual(labels= c("P25", "P75"), values=c("blue","red"))+
  theme_minimal()

dev.off()


#DID

birthall$time = ifelse(birthall$year>1996,1,0)
birthall$treated = ifelse(birthall$percent_change>=x.75th,1,0)
birthall$did = birthall$time * birthall$treated

didreg = lm(mx_mean ~ treated + time + did, dat = birthall)

sink(paste0("R:/working/preparing data/Density/Figures/didsummary_75v25_", this_cause, ".txt")) 

print(summary(didreg)) 

sink()

```

## Apply DID model on cause-specific mortalities - low SES group


```{r}
#this_cause<-"cvd"
dir2<-"R:/working/preparing data/"
file<-".Rdata"
load(paste0(dir2, "density/birthAll_",this_cause, file))

load("R:/working/preparing data/density/densityChangeCheck.RDATA")

birthAll3<-birthAll3 %>% filter(msdgroup==1) #1 is low in this dataset

birthAll3<-merge(birthAll3,densityChangeCheck,by="area")

birthall<-birthAll3


x.90th<-quantile(birthall$percent_change, c(.90))
x.75th<-quantile(birthall$percent_change, c(.75))


birthall<-subset(birthall, percent_change>=quantile(birthall$percent_change, c(.90)) | percent_change<=quantile(birthall$percent_change, c(.1)))

#Find density change data 
birthall = birthall %>%
  mutate(treated = percent_change>=x.90th)

#Plot year and life expectancy of two density change categories to ensure parallel assumption is met
pdf(paste0("R:/working/preparing data/Density/Figures/parallel_line_assumption_90v10_", this_cause, "_lowSES.pdf")) 

ggplot(birthall, aes(year, mx_mean, color = treated)) +
  stat_summary(geom = 'line') +
  geom_vline(xintercept = 1996) +
  ggtitle("Parallel Line Assumption for Densification of P90 vs P10 - All CTs")+
  xlab("Year") +
  ylab("Life Expectancy") +
  labs(color="Densification") +
  scale_color_manual(labels= c("P10", "P90"), values=c("blue","red"))+
  theme_minimal()

dev.off()


#DID

birthall$time = ifelse(birthall$year>1996,1,0)
birthall$treated = ifelse(birthall$percent_change>=x.90th,1,0)
birthall$did = birthall$time * birthall$treated

didreg = lm(mx_mean ~ treated + time + did, dat = birthall)
CIs<-confint(didreg, "did", level=0.95)
assign(paste0("CIs_",this_cause, "_low"), CIs)


sink(paste0("R:/working/preparing data/Density/Figures/didsummary_90v10_", this_cause, "_lowSES.txt")) 

print(summary(didreg)) 

sink()

#75 vs 25
birthall<-birthAll3

birthall<-subset(birthall, percent_change>=quantile(birthall$percent_change, c(.75)) | percent_change<=quantile(birthall$percent_change, c(.25)))

#Find density change data 
birthall = birthall %>%
  mutate(treated = percent_change>=x.75th)

#Plot year and life expectancy of two density change categories to ensure parallel assumption is met
pdf(paste0("R:/working/preparing data/Density/Figures/parallel_line_assumption_75v25_", this_cause, "_lowSES.pdf")) 

ggplot(birthall, aes(year, mx_mean, color = treated)) +
  stat_summary(geom = 'line') +
  geom_vline(xintercept = 1996) +
  ggtitle("Parallel Line Assumption for Densification of P75 vs P25 - All CTs")+
  xlab("Year") +
  ylab("Life Expectancy") +
  labs(color="Densification") +
  scale_color_manual(labels= c("P25", "P75"), values=c("blue","red"))+
  theme_minimal()

dev.off()


#DID

birthall$time = ifelse(birthall$year>1996,1,0)
birthall$treated = ifelse(birthall$percent_change>=x.75th,1,0)
birthall$did = birthall$time * birthall$treated

didreg = lm(mx_mean ~ treated + time + did, dat = birthall)

sink(paste0("R:/working/preparing data/Density/Figures/didsummary_75v25_", this_cause, "_lowSES.txt")) 

print(summary(didreg)) 

sink()

```


## Apply DID model on cause-specific mortalities - high SES group


```{r}
#this_cause<-"mater_neonat"
dir2<-"R:/working/preparing data/"
file<-".Rdata"
load(paste0(dir2, "density/birthAll_",this_cause, file))

load("R:/working/preparing data/density/densityChangeCheck.RDATA")

birthAll3<-birthAll3 %>% filter(msdgroup==5)

birthAll3<-merge(birthAll3,densityChangeCheck,by="area")

birthall<-birthAll3


x.90th<-quantile(birthall$percent_change, c(.90))
x.75th<-quantile(birthall$percent_change, c(.75))


birthall<-subset(birthall, change>=quantile(birthall$percent_change, c(.90)) | change<=quantile(birthall$percent_change, c(.1)))

#Find density change data 
birthall = birthall %>%
  mutate(treated = percent_change>=x.90th)

#Plot year and life expectancy of two density change categories to ensure parallel assumption is met
pdf(paste0("R:/working/preparing data/Density/Figures/parallel_line_assumption_90v10_", this_cause, "_highSES.pdf")) 

ggplot(birthall, aes(year, mx_mean, color = treated)) +
  stat_summary(geom = 'line') +
  geom_vline(xintercept = 1996) +
  ggtitle("Parallel Line Assumption for Densification of P90 vs P10 - All CTs")+
  xlab("Year") +
  ylab("Life Expectancy") +
  labs(color="Densification") +
  scale_color_manual(labels= c("P10", "P90"), values=c("blue","red"))+
  theme_minimal()

dev.off()


#DID

birthall$time = ifelse(birthall$year>1996,1,0)
birthall$treated = ifelse(birthall$percent_change>=x.90th,1,0)
birthall$did = birthall$time * birthall$treated

didreg = lm(mx_mean ~ treated + time + did, dat = birthall)
CIs<-confint(didreg, "did", level=0.95)
assign(paste0("CIs_",this_cause,"_high"), CIs)


sink(paste0("R:/working/preparing data/Density/Figures/didsummary_90v10_", this_cause, "_highSES.txt")) 

print(summary(didreg)) 

sink()

#75 vs 25

birthall<-birthAll3

birthall<-subset(birthall, change>=quantile(birthall$percent_change, c(.75)) | change<=quantile(birthall$percent_change, c(.25)))

#Find density change data 
birthall = birthall %>%
  mutate(treated = percent_change>=x.75th)

#Plot year and life expectancy of two density change categories to ensure parallel assumption is met
pdf(paste0("R:/working/preparing data/Density/Figures/parallel_line_assumption_75v25_", this_cause, "_highSES.pdf")) 

ggplot(birthall, aes(year, mx_mean, color = treated)) +
  stat_summary(geom = 'line') +
  geom_vline(xintercept = 1996) +
  ggtitle("Parallel Line Assumption for Densification of P75 vs P25 - All CTs")+
  xlab("Year") +
  ylab("Life Expectancy") +
  labs(color="Densification") +
  scale_color_manual(labels= c("P25", "P75"), values=c("blue","red"))+
  theme_minimal()

dev.off()


#DID

birthall$time = ifelse(birthall$year>1996,1,0)
birthall$treated = ifelse(birthall$percent_change>=x.75th,1,0)
birthall$did = birthall$time * birthall$treated

didreg = lm(mx_mean ~ treated + time + did, dat = birthall)

sink(paste0("R:/working/preparing data/Density/Figures/didsummary_75v25_", this_cause, "_highSES.txt")) 

print(summary(didreg)) 

sink()

```


## Descriptive data for paper

```{r}
CTs<-unique(birthall$area)

allCTs<-birthall %>% filter(year<1996) %>% summarise(ex_mean=median(ex_mean), population_density=median(population_density), msd=median(msd), education=median(education), aboriginal=median(Aboriginal))

treatedCTs<-birthall %>% filter(treated==1) %>% group_by(area) %>% summarise(CSDNAME=max(CSDNAME))
controlCTs<-birthall %>% filter(treated==0) %>% group_by(area) %>% summarise(CSDNAME=max(CSDNAME))



treated<-birthall %>% filter(treated==1 & year<1996) %>% summarise(ex_mean=median(ex_mean), population_density=median(population_density), msd=median(msd), education=median(education), aboriginal=median(Aboriginal), MatScore=median(MatScore), SocScore=median(SocScore))

control<-birthall %>% filter(treated==0 & year<1996) %>%  summarise(ex_mean=median(ex_mean),population_density=median(population_density), msd=median(msd), education=median(education), aboriginal=median(Aboriginal), MatScore=median(MatScore), SocScore=median(SocScore))

```

## Derive land area

```{r}
data<-birthall
load("R:/working/preparing data/CT368.rdata")

CTs<-data.frame(area=unique(data$area))
CTs<-left_join(CTs, CT, by="area")

landarea<-left_join(CTs, densitycwed, by="CT") %>%  summarise(land_area=median(land_area))

```

